### 2018-10-03 raw_notes_questions.rmd
Update raw_notes_questions.rmd

# Question
## When I write to the Master, I am writing to the forked copy of this website, correct? 
So if that is correct, and Mike is still listed as a collaborator:
Are MY edits being kept here until I send a Pull/Push request?
My edits only affect this repo, correct?