### 2018-10-03 raw_notes_questions.rmd
Update

# Question
#### When I write to the Master, I am writing to my forked copy of the website, correct? 
If Mike is still listed as a collaborator, MY edits are only pushed to my version in my repo, and affect only my copy. If I send a Push request, Mike can pull in my changes IF HE CHOOSES?

# Question
#### Anyone else need a QUICK tour of the branch and pull request map?

# Question
#### Still not sure I know difference between a fork and a branch. 
Branch is an alternative version of a single file, but fork is a copy of the original repo in total. Y/N?

# Question
#### Can I build multiple files in R, then push project to GitHub as a complete repo?
Yep. Found instructions for doing it on this website.
https://cfss.uchicago.edu/git05.html

Is this correct, that if need be, I could download a ZIP of a project from someone else, unpack in my local machine (Dropbox, actually), then upload back to GitHub just the same as if I cloned it directly inside GH?

I know it is the long way around, but if I have a folder of resources someone gives me, it seems rude to ask, "can you put that on GitHub first for me?"

So how do I tell R to use my GitHub repo's as their source for data?

# Question
#### Having a slight problem with Git on laptop.
I can use my desktop and create repos to Dropbox or local machine from GitHub. Laptop throws in an extra step that seems to corrupt process.

**SOLVED**. The path to Git was wrong. Needs to be very specific: 
/usr/local/git/bin/git

Other Git and Git-GUI paths do not work.