### 2018-10-31 project.Rmd
update

---
title: "Project"
author: "Dan Johnson"
date: "10/26/2018"
runtime: shiny
output: html_document
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```

# Abstract

### Here are two Shiny widgets

```{r echo = FALSE}
selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20)

sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
```

### ...that build a histogram.

```{r echo = FALSE}
renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")

  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

# Introduction

# Literature Review

# Methodology

## Data

## Analysis

# Results

# Discussion

# Conclusion

# References