### 2018-11-22 course_notes.Rmd
change

---
title: "Course Notes"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

You can detail what you learned in the session within these pages. You can also use the Rmd markdown language and insert code chunks with `CMD/CTRL + SHIFT + I`.

# Session 1
Change

# Session 2

# Session 3

# Session 4

# Session 5

# Session 6

# Session 7

# Session 8

# References