### 2018-11-29 course_notes.Rmd
updates

---
title: "Notes & Ideas"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

# Session 1 (Oct. 1, 2018)
Change

####\

# Session 2 (Oct. 22, 2018)

####\

# Session 3 (Nov. 12, 2018)

####\

# Session 4 (Dec. 3, 2018)

####\

# Session 5 (Jan. 14, 2019)

####\

# Session 6 (Feb. 4, 2019)

####\

# Session 7 (Feb. 25, 2019)

####\

# Session 8 (Mar. 18, 2019)

####\

# References

####\