### 2018-11-30 questions.rmd

# R Code
#### I cannot get numbers on CSV extracted data frames.

Ideas? Detailed notes are in "Data_intake_1.rmd", starting Line 102.

#### Why can I not append the read.csv command?
This works fine:
```{r}
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18.csv')
head(base_data)
```

This fails:
```{r}
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18.csv', header=FALSE)
head(base_data)
```
I get:
```	
	Error in read_csv(file = "data/coded_full_comments_dataset_Spring18.csv",  : 
  unused argument (header = FALSE)
```

-----

# Project

-----

# Github
#### Question about fork and branch. 
When I fork another repo, I get comments saying I am "#" steps ahead of it. Should I clone instead?

-----

#### Can I build multiple files in R, then push project to GitHub as a complete repo?
Yep. Found instructions for doing it on this website.
https://cfss.uchicago.edu/git05.html

Is this correct, that if need be, I could download a ZIP of a project from someone else, unpack in my local machine (Dropbox, actually), then upload back to GitHub just the same as if I cloned it directly inside GH?

I know it is the long way around, but if I have a folder of resources someone gives me, it seems rude to ask, "can you put that on GitHub first for me?"

So how do I tell R to use my GitHub repo's as their source for data?

-----

####\