### 2019-01-31 Naive_bayes_V2r.Rmd
Working text analysis forms

---
title: "Naive Bayes Workflow Ver 2.0.r"
author: "Dan Johnson"
date: "1/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Summary
This is **Version 2.0.r** of the basic workflow for classifying TA comments using a naive Bayes model. It pre-processes text using the same command set as Ver 2.0.b. The "r" designates this version has been modified to be able to randomize which rows are selected for training and test datasets.

####\

##Initial Setup
Calling the required libraries.

```{}
library(tidyverse)
library(tidytext)
library(tidyr)
library(tm)
library(SnowballC)
library(wordcloud)
library(e1071)
library(gmodels)
```

####\

##Version 2.0.r of Naive Bayes (NB) Protocol
###Initial Data Input
```{r}
#Read in TA comments from CSV file.
base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

###Extract, Organize Data Subset
Extract comments and classification data subset
```{r}
#Select rows representing the sub-groups to compare. 
comments_subsetVERB <- filter(base_dataVERB,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")

#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
comments_rawVERB <- comments_subsetVERB %>% select(23,24,22)

#Rename the columns.
names(comments_rawVERB)[1] <- "subject"
names(comments_rawVERB)[2] <- "structure"
names(comments_rawVERB)[3] <- "text"

#Change "subject" element from character to a factor for analysis.
comments_rawVERB$subject <- factor(comments_rawVERB$subject)
str(comments_rawVERB$subject)
table(comments_rawVERB$subject)
```

```{r}
#Change "structure" element from character to a factor for analysis.
comments_rawVERB$structure <- factor(comments_rawVERB$structure)
str(comments_rawVERB$structure)
table(comments_rawVERB$structure)
```

```{r}
#Generate a vector of ~10,000 numbers (be sure it can be generated consistently)
#Adjust the number so it matches the number of rows in the original data frame
#Next, set the seed for random numbers, and randomize the order of the vector
vector <- 1:9340
set.seed(123) #change this seed value, and the data are randomized in a new sequence.
vector <- sample(vector)
glimpse(vector)
```

```{r}
#Append the vector as a new column to "comments_rawVERB" with name "randomizer"
#Sort rows of dataframe according to "randomizer" column using dplyer 'arrange' function. 
comments_rawVERB$randomizer <- c(vector)
comments_rawVERB_R <-comments_rawVERB %>% arrange(randomizer)
print(comments_rawVERB_R)

```

Convert data set to a volative corpus using "tm" library.
```{r}
#Create the volatile corpus that contains the "text" vector from data frame.
comments_corpusVERB_R <- VCorpus(VectorSource(comments_rawVERB_R$text))
print(comments_corpusVERB_R)

#Check out the first few text comments in the new corpus, which is basically a list that can be manipulated with list operations.
inspect(comments_corpusVERB_R[1:3])

#Use "as.character" function to see what a single text comment looks like.
as.character(comments_corpusVERB_R[[5]])
```

###Text Data Transformations
The text data transforms are separated out as individual "tm_map"" and "content_transformer" functions to simplify the process of testing various combinations. Conversion to lower case and removing punctuation are the baseline set. Switch between "{r}" and "{}" to turn the other transformation blocks off or on. Removing extra white spaces must be the last transformation, and is required.

```{r}
#Convert text to all lower case letters.
comments_corpus_cleanVERB_R <- tm_map(comments_corpusVERB_R, content_transformer(tolower))

#Remove punctuation using the "removePunctuation" function. 
#This step removes evidence of questions, so may remove data.
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removePunctuation)

#Remove numerals. 
#I am not sure this is beneficial; does it remove evidence indicating technical comments?
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removeNumbers)

#Stopword removal. 
#This is a standard cleanup step, but again may remove useful terms.
#The "stopwords" file is a convention; other files can be substituted for it
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removeWords, stopwords())

#The option to use stemming has been removed from this version.

#Final step removes white space from the document. This is NOT an optional step.
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, stripWhitespace)

#Look at an example of cleaned text comment to see if it fits what is expected.
as.character((comments_corpus_cleanVERB_R[[5]]))
```

####\

##Basic Naive Bayes (NB) Method Using N-grams
This code block uses tm's NLP commands. The same block can be used to assess 1-, 2-, or 3-grams. Only the "#" variable in "(words(x), #)" below needs to be changed.

```{r}
NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 3), paste, collapse = " "), use.names = FALSE)
}

comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVERB_R, control=list(tokenize = NLP_Tokenizer))
comments_dtm_1gram_tidy <- tidy(comments_dtm_1gram)
inspect(comments_dtm_1gram)
```

####\

###Data Preparation
Split data into 75% training and 25% testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data.

```{}
.75 * 9340 #Number of rows in the dataset; product is #rows for training
.25 * 9340 #Product is #rows for testing set
```

This code uses the comments in the order they are listed in the original file. Block for randomizing them was done earlier.
```{r}
comments_dtm_train <- comments_dtm_1gram[1:7005, ]
comments_dtm_test <- comments_dtm_1gram[7006:9340, ]
```

Save vectors labeling rows in the training and testing vectors
```{r}
comments_train_labels_subject <- comments_rawVERB_R[1:7005, ]$subject
comments_test_labels_subject <- comments_rawVERB_R[7006:9340,]$subject
```

```{r}
comments_train_labels_structure <- comments_rawVERB_R[1:7005, ]$structure
comments_test_labels_structure <- comments_rawVERB_R[7006:9340,]$structure
```

Make sure that the proportion of each sub-category is similar in the training and testing data sets.
```{r}
prop.table(table(comments_train_labels_subject))
prop.table(table(comments_test_labels_subject))
```

```{r}
prop.table(table(comments_train_labels_structure))
prop.table(table(comments_test_labels_structure))
```

####\

###Preparation for Naive Bayes
Remove words from the matrix that appear less than 5 times.
```{r}
comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
str(comments_freq_words)
```

Limit Document Term Matrix to only include words in the comments_freq_words vector. Using all the rows, but we want to limit the columns to these words in the frequency vector.
```{r}
comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
```

The basic e1071 naive bayes classifier works with categorical features, so the matrix must be converted "yes" and "no" categorical variables. This is done using a **convert_counts function** and applying it to the data. This replaces values greater than 0 with yes, and values not greater than 0 with no. 
```{r}
convert_counts2 <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}

comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
```

The resulting matrices have cells indicating "yes" or "no" if the word represented by the column appears in the text comment represented by the row.

####\

###Train Model, Predict, Evaluate for Subject
This block uses the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. It evaluates the prediction with the actual data using a crosstable from the gmodels package.
```{r}
comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject)
comments_test_pred_subject <- predict(comments_classifier_subject, comments_test)
CrossTable(comments_test_pred_subject, comments_test_labels_subject, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

The cross table lists how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting subject.
```{r}
print("Overall subject accuracy")
(14+54+1318+28)/23.35  #Need to find a way to calculate this automatically. 
```

###Train Model, Predict, Evaluate for structure
This is the same block as above but uses the "structure" data column.

```{r}
comments_classifier_structure <- naiveBayes(comments_train, comments_train_labels_structure)
comments_test_pred_structure <- predict(comments_classifier_structure, comments_test)
CrossTable(comments_test_pred_structure, comments_test_labels_structure, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

Again, a cross table shows how many of the predicted categories match the correct categories for structure.

```{r}
print("Overall structure accuracy")
(66+64+949+5+1)/23.35  #Need to find a way to calculate this automatically. 
```

###OPTIONAL Output
To make a table showing predicted categories for comments based on the Naive Bayes test, versus the actual label entered by a human observer, merge the two datasets into a single data frame. The full data frame will be too long to display; use "head" to check format, then output the full dataset to CSV for further review. By default the CSV file should be saved to the current working directory.

```{}
predicted_category_subject <- data.frame(comments_test_pred_subject,comments_test_labels_subject)
head(predicted_category_subject)
write.csv(predicted_category_subject, file = "PredictedCategorySubject_Data.csv")
```

```{}
predicted_category_structure <- data.frame(comments_test_pred_structure,comments_test_labels_structure)
head(predicted_category_structure)
write.csv(predicted_category_structure, file = "PredictedCategoryStructure_Data.csv")
```

***
START COMMENTS, OLD work notes
***

##Permutation Tests of Naive Bayes (NB) Method
Using single words, Version 2.0.b correctly identifies comment "subject" ~80% of the time and comment "structure" ~59% of the time.

Permutations to test with random subsets version (2.0.r):

*  Compare 1-grams, 2-grams, 3-grams
*  Compare outcomes for 3 independent randomized sets of 1-grams

The pseudo-code block below can pull random lines from a data table, but does not work correctly on a document term matrix. The second problem is how to pull the training and test data sets out and process them identically.

```{}
set.seed(1234) 		# ensures the same random set gets picked each time.

comments_raw_dtm_train <- sample_frac(comments_raw, .75)	# sample_frac pulls out 75% of "comments_raw" dataset randomly using the sample sequence identified by set.seed(1234)

comments_raw_dtm_test <- setdiff(comments_raw, comments_raw_dtm_train)	#setdiff creates a "test" dataset using anything NOT in the "comments_raw_dtm_train" subset of "comments_raw."
```

The solution was not to pull the data twice, but to randomly (but repeatably) shuffle the starting data table. How I solved this was:
1. Create a vector of sequential numbers the same length as the number of data frame rows.
2. Randomize the sequential numbers using set.seed to make pattern reproducible.
3. Add the randomized numbers as a new column to the data frame.
4. Re-order the data table to the randomizer column.

This makes it easy to generate a new sequence of randomized training and test samples to run replicates. Only one number needs to be changed.

```{}
#Creates a data frame that compares predicted and actual labels for subject and structure
tally <- data.frame("subject_label"=c(comments_test_labels_subject), "subject_predicted"=c(comments_test_pred_subject), "structure_label"=c(comments_test_labels_structure), "structure_predicted"=c(comments_test_pred_structure))
#tally <- as.numeric(as.character(tally))
```

```{}
if_else (tally(comments_test_labels_subject) != tally(comments_test_pred_subject),"YES","no")

#glimpse(tally)

#tally %>% mutate(match_subject=comments_test_labels_subject - comments_test_pred_subject)
#tally %>% mutate(match_structure=comments_test_labels_structure - comments_test_pred_structure)
```

***
END COMMENTS
***