### 2019-01-31 Naive_bayes_V2yn.Rmd

---
title: "Naive Bayes Workflow Ver 2.0.yn"
author: "Dan Johnson"
date: "1/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Summary
This is **Version 2.0.yn** of the basic workflow for classifying TA comments using a naive Bayes model. It is optimized for testing pair-wise assignment to categories (Yes/No decision) beginning with the most frequently encountered.

The main features and permutations to be evaluated are:

*  Stepwise paired classification versus comparing multiple groups simultaneously

Additional working notes are located at the end of this page.

####\

##Initial Setup
Calling the required libraries.

```{r}
library(tidyverse)
library(tidytext)
library(tidyr)
library(tm)
library(SnowballC)
library(wordcloud)
library(e1071)
library(gmodels)
```

####\

##Version 2.0.yn of Naive Bayes (NB) Protocol
###Initial Data Input
```{r}
#Read in TA comments from CSV file.
base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

###Extract, Organize Data Subset
Extract comments and classification data subset
```{r}
#Select rows representing the sub-groups to compare. 
comments_subsetVERB <- filter(base_dataVERB,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")

#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
comments_rawVERB <- comments_subsetVERB %>% select(23,24,22)

#Rename the columns.
names(comments_rawVERB)[1] <- "type"
names(comments_rawVERB)[2] <- "locus"
names(comments_rawVERB)[3] <- "text"

#Change "type" element from character to a factor for analysis.
comments_rawVERB$type <- factor(comments_rawVERB$type)
str(comments_rawVERB$type)
table(comments_rawVERB$type)
```

```{r}
#Change "locus" element from character to a factor for analysis.
comments_rawVERB$locus <- factor(comments_rawVERB$locus)
str(comments_rawVERB$locus)
table(comments_rawVERB$locus)
```

***
***
Here is where I need to add the column used to randomly assign the data to training or testing sets.

Use set.seed(1234) to create a stable randomized set.

***
***

Convert data set to a volative corpus using "tm" library.
```{r}
#Create the volatile corpus that contains the "text" vector from data frame.
comments_corpusVERB <- VCorpus(VectorSource(comments_rawVERB$text))
print(comments_corpusVERB)

#Check out the first few text comments in the new corpus, which is basically a list that can be manipulated with list operations.
inspect(comments_corpusVERB[1:3])

#Use "as.character" function to see what a single text comment looks like.
as.character(comments_corpusVERB[[5]])
```

###Text Data Transformations
The text data transforms are separated out as individual "tm_map"" and "content_transformer" functions to simplify the process of testing various combinations. Conversion to lower case and removing punctuation are the baseline set. Switch between "{r}" and "{}" to turn the other transformation blocks off or on. Removing extra white spaces must be the last transformation, and is required.

```{r}
#Convert text to all lower case letters.
comments_corpus_cleanVERB <- tm_map(comments_corpusVERB, content_transformer(tolower))
```

```{r}
#Remove punctuation using the "removePunctuation" function. 
#This step removes evidence of questions, so may remove data.
comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, removePunctuation)
```

```{r}
#Remove numerals. 
#I am not sure this is beneficial; does it remove evidence indicating technical comments?

comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, removeNumbers)
```

```{r}
#Stopword removal. 
#This is a standard cleanup step, but again may remove useful terms.
#The "stopwords" file is a convention; other files can be substituted for it

comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, removeWords, stopwords())
```

Option to stem terms has been removed from this version.

Final prep and data check.
```{r}
#Final step removes white space from the document. This is NOT an optional step.
comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, stripWhitespace)

#Look at an example of cleaned text comment to see if it fits what is expected.
as.character((comments_corpus_cleanVERB[[5]]))
```

####\

##Basic Naive Bayes (NB) Method Using N-grams
This code block uses tm's NLP commands. The same block can be used to assess 1-, 2-, or 3-grams. Only the "#" variable in "(words(x), #)" below needs to be changed.

```{r}
NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
}

comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVERB, control=list(tokenize = NLP_Tokenizer))
comments_dtm_1gram_tidy <- tidy(comments_dtm_1gram)
inspect(comments_dtm_1gram)
```

####\

###Data Preparation
Split data into 75% training and 25% testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data.

```{}
.75 * 9340 #Number of rows in the dataset; product is #rows for training
.25 * 9340 #Product is #rows for testing set
```

This code uses the comments in the order they are listed in the original file. **Need a block for  randomizing them.**
```{}
comments_dtm_train <- comments_dtm_1gram[1:7005, ]
comments_dtm_test <- comments_dtm_1gram[7006:9340, ]
```

Save vectors labeling rows in the training and testing vectors
```{}
comments_train_labels_type <- comments_rawVERB[1:7005, ]$type
comments_test_labels_type <- comments_rawVERB[7006:9340,]$type
```

```{}
comments_train_labels_locus <- comments_rawVERB[1:7005, ]$locus
comments_test_labels_locus <- comments_rawVERB[7006:9340,]$locus
```

Make sure that the proportion of each sub-category is similar in the training and testing data set.
```{}
prop.table(table(comments_train_labels_type))
prop.table(table(comments_test_labels_type))
```

```{}
prop.table(table(comments_train_labels_locus))
prop.table(table(comments_test_labels_locus))
```

####\

###Preparation for Naive Bayes
Remove words from the matrix that appear less than 5 times.
```{r}
comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
str(comments_freq_words)
```

Limit Document Term Matrix to only include words in the comments_freq_words vector. Using all the rows, but we want to limit the columns to these words in the frequency vector.
```{r}
comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
```

The basic e1071 naive bayes classifier works with categorical features, so the matrix must be converted "yes" and "no" categorical variables. This is done using a **convert_counts function** and applying it to the data. This replaces values greater than 0 with yes, and values not greater than 0 with no.

```{r}
convert_counts2 <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}

comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
```

The resulting matrices have cells indicating "yes" or "no" if the word represented by the column appears in the text comment represented by the row.

####\

###Train Model, Predict, Evaluate for Type
This block uses the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. It evaluates the prediction with the actual data using a crosstable from the gmodels package.
```{r}
comments_classifier_type <- naiveBayes(comments_train, comments_train_labels_type)
comments_test_pred_type <- predict(comments_classifier_type, comments_test)
CrossTable(comments_test_pred_type, comments_test_labels_type, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

The cross table list how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting TYPE.

```{r}
print("Overall TYPE accuracy")
(14+45+1346+19)/23.35
```

###Train Model, Predict, Evaluate for LOCUS
This is the same block as above but uses the "Locus" data column.

```{r}
comments_classifier_locus <- naiveBayes(comments_train, comments_train_labels_locus)
comments_test_pred_locus <- predict(comments_classifier_locus, comments_test)
CrossTable(comments_test_pred_locus, comments_test_labels_locus, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

Again, a cross table shows how many of the predicted categories match the correct categories for LOCUS.

```{r}
print("Overall LOCUS accuracy")
(62+64+931+3)/23.35
```

###OPTIONAL Output
To make a table showing predicted categories for comments based on the Naive Bayes test, versus the actual label entered by a human observer, merge the two datasets into a single data frame. The full data frame will be too long to display; use "head" to check format, then output the full dataset to CSV for further review. By default the CSV file should be saved to the current working directory.

```{}
predicted_category_type <- data.frame(comments_test_pred_type,comments_test_labels_type)
head(predicted_category_type)
write.csv(predicted_category_type, file = "PredictedCategoryType_Data.csv")
```

```{}
predicted_category_locus <- data.frame(comments_test_pred_locus,comments_test_labels_locus)
head(predicted_category_locus)
write.csv(predicted_category_locus, file = "PredictedCategoryLocus_Data.csv")
```

***
START COMMENTS
***

##Permutation Tests of Naive Bayes (NB) Method
Using single words, Version 2.0.b correctly identifies comment "type" ~80% of the time and comment "locus" ~59% of the time.

Permutations to Test:

*  Compare 1-grams, 2-grams, 3-grams
*  Make pair-wise comparisons versus comparing multiple groups simultaneously. 
    +  Initial approach will be to split between largest group and balance in each category
    +  Ex.: type is 3-technical vs. not 3_technical

Known Issues:

1. The e1071 naive bayes classifier is binomial. Need to incorporate multinomial.

***
END COMMENTS
***