### 2019-02-05 Backups/Naive_bayes_V2a.Rmd
Update

---
title: "Naive Bayes Workflow Ver 2.0"
author: "Dan Johnson"
date: "1/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Summary
This notebook documents a basic workflow for applying a naive Bayes model to terms in TA comments. I based it on a tutorial [https://rpubs.com/Billyhansen6/318412](located here). This is quoted from Hansen's original tutorial:
>The objective of this project is to classify SMS messages as spam or ham (not spam). A Naive Bayes classifier approach will be used. This example is taken from chapter 4 of Machine Learning with R, Second Edition"

>An example of the conditional probability that will be computed is as follows:

$$P(Spam|Hospital) = P(Hospital|Spam)P(Spam)/P(Hospital)$$

>which is the formula for determining the probability that a message is spam given that it contains the word "Hospital" in the message.

Lines ~30-205 are the baseline version of the workflow using single words. It correctly classifies comments in the top 4 sub-categories  ~79% of the time. While that is close to minimum acceptable inter-rater reliability of two human readers, it is well below my goal of >95% for use in routine automated analysis.

Lines ~208-290 describe an alternative method using n-grams. When I tested this approach using 2-grams, classification was slightly worse, at 68.3% correct.

Lines ~295-335 show how to use word clouds to visualize single word frequencies. I did not find this approach very useful.

To keep this code clean for now I will run various permutations in a scratch version, then summarize findings starting on line #345. In the subsequent iterations I will evaluate:

*  Different text pre-processing strategies: stemmed versus unstemmed, numbers removed vs. not, etc.
*  Using n-grams rather than single words
*  Making pair-wise comparison versus comparing multiple groups simultaneously
*  Modifying the variables in the NB analysis itself.

Finally, there are several things still missing or needed for this workflow:

*  How to feed a text dataset in, and get a table of the predicted categories back out.
*  How to store the values from the Bayes prediction table so I can use them in calculations rather than type them in by hand.
*  How to randomize which comments are used for training, testing datasets.

Other pages examine other classifiers besides Naive Bayes.

####\

##Initial Setup
```{r}
library(tidyverse)
library(tidytext)
library(tidyr)
library(tm)
library(SnowballC)
library(wordcloud)
library(e1071)
library(gmodels)
```

####\

##Version 2.0 of Naive Bayes (NB) Protocol
###Initial Data Input
```{r}
#Read in TA comments from CSV file.
base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

###Extract, Organize Data Subset
Extract comments and classification data subset
```{r}
#Select rows representing the sub-groups to compare. 
comments_subsetVERB <- filter(base_dataVERB,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")

#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
comments_rawVERB <- comments_subsetVERB %>% select(23,24,22)

#Rename the columns.
names(comments_rawVERB)[1] <- "type"
names(comments_rawVERB)[2] <- "locus"
names(comments_rawVERB)[3] <- "text"

#Change "type" element from character to a factor for analysis.
comments_rawVERB$type <- factor(comments_rawVERB$type)
str(comments_rawVERB$type)
table(comments_rawVERB$type)
```
```{r}
#Change "locus" element from character to a factor for analysis.
comments_rawVERB$locus <- factor(comments_rawVERB$locus)
str(comments_rawVERB$locus)
table(comments_rawVERB$locus)
```

***
***
Here is where I need to add the column used to randomly assign the data to training or testing sets.

Use set.seed(1234) to create a stable randomized set.

***
***

Convert data set to a volative corpus using "tm" library.
```{r}
#Create the volatile corpus that contains the "text" vector from data frame.
comments_corpusVERB <- VCorpus(VectorSource(comments_rawVERB$text))
print(comments_corpusVERB)

#Check out the first few text comments in the new corpus, which is basically a list that can be manipulated with list operations.
inspect(comments_corpusVERB[1:3])

#Use "as.character" function to see what a single text comment looks like.
as.character(comments_corpusVERB[[5]])
```

###Text Data Transformations
The text data transforms are separated out as individual "tm_map"" and "content_transformer" functions to simplify the process of testing various combinations. Conversion to lower case and removing punctuation are the baseline set. Switch between "{r}" and "{}" to turn the other transformation blocks off or on. Removing extra white spaces must be the last transformation, and is required.

```{r}
#Convert text to all lower case letters.
comments_corpus_cleanVERB <- tm_map(comments_corpusVERB, content_transformer(tolower))
```

```{r}
#Remove punctuation using the "removePunctuation" function. 
#This step removes evidence of questions, so may remove data.
comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, removePunctuation)
```

```{}
#Remove numerals. 
#I am not sure this is beneficial; does it remove evidence indicating technical comments?

comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, removeNumbers)
```

```{}
#Stopword removal. 
#This is a standard cleanup step, but again may remove useful terms.
#The "stopwords" file is a convention; other files can be substituted for it

comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, removeWords, stopwords())
```

***
START BREAK POINT - Stemming vs. lemmatization
***

```{}
#Stemming the text data to strip the suffix from words. 
#My thinking is that stemming removes valuable tense data.I am unsure how valuable it is.
#Looking to later versions, part-of-speech tagging and lemmatization may be the better method.

comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, stemDocument)
```

***
END BREAK POINT
***

Final prep and data check.
```{r}
#Final step removes white space from the document. This is NOT an optional step.
comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, stripWhitespace)

#Look at an example of cleaned text comment to see if it fits what is expected.
as.character((comments_corpus_cleanVERB[[5]]))
```

####\

##Basic Naive Bayes (NB) Method Using N-grams
This code block uses tm's NLP commands. The same block can be used to assess 1-, 2-, or 3-grams. Only the "#" variable in "(words(x), #)" below needs to be changed.

```{r}
NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
}

comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVERB, control=list(tokenize = NLP_Tokenizer))
comments_dtm_1gram_tidy <- tidy(comments_dtm_1gram)
inspect(comments_dtm_1gram)
```

####\

###Data Preparation
Split data into 75% training and 25% testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data.

```{r}
.75 * 9340 #Number of rows in the dataset; product is #rows for training
.25 * 9340 #Product is #rows for testing set
```

This code assumes comments are random. **Probably want to try randomizing them.**
```{r}
comments_dtm_train <- comments_dtm_1gram[1:7005, ]
comments_dtm_test <- comments_dtm_1gram[7006:9340, ]
```

Save vectors labeling rows in the training and testing vectors
```{r}
comments_train_labels <- comments_rawVERB[1:7005, ]$type
comments_test_labels <- comments_rawVERB[7006:9340,]$type
```

```{r}
comments_train_labels_locus <- comments_rawVERB[1:7005, ]$locus
comments_test_labels_locus <- comments_rawVERB[7006:9340,]$locus
```

Make sure that the proportion of each sub-category is similar in the training and testing data set.
```{r}
prop.table(table(comments_train_labels))
prop.table(table(comments_test_labels))
```

```{r}
prop.table(table(comments_train_labels_locus))
prop.table(table(comments_test_labels_locus))
```

***
START BREAK POINT
***
The above code block incorrectly uses sequential roww, not random rows.  I need to find a way to incorporate randomization.

The command block below can pull random lines from a data table, but does not work on a document term matrix.

```{}
set.seed(1234) 		# ensures the same random set gets picked each time.

comments_raw_dtm_train <- sample_frac(comments_raw, .75)	# sample_frac pulls out 75% of "comments_raw" dataset randomly using the sample sequence identified by set.seed(1234)

comments_raw_dtm_test <- setdiff(comments_raw, comments_raw_dtm_train)	#setdiff creates a "test" dataset using anything NOT in the "comments_raw_dtm_train" subset of "comments_raw."
```

Ways I think will work to solve this:
1. Embed a randomized identifier as another column, then pull the later datasets out using that identifier.
2. Generate a vector of random words (test, train, test, train, train, train, test, train...)

***
END BREAK POINT
***

####\

###Preparation for Naive Bayes
Remove words from the matrix that appear less than 5 times.
```{r}
comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
str(comments_freq_words)
```

Limit our Document Term Matrix to only include words in the sms_freq_vector. We want all the rows, but we want to limit the column to these words in the frequency vector.
```{r}
comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
```

The basic e1071 naive bayes classifier works with categorical features, so the matrix must be converted "yes" and "no" categorical variables. This is done using a **convert_counts function** and applying it to the data. This replaces values greater than 0 with yes, and values not greater than 0 with no.

```{r}
convert_counts2 <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}

comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
```

The resulting matrices have cells indicating "yes" or "no" if the word represented by the column appears in the text comment represented by the row.

####\

###Train Model, Predict, Evaluate
Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. Evaluate the prediction with the actual data using a crosstable from the gmodels package.
```{r}
comments_classifier <- naiveBayes(comments_train, comments_train_labels)
comments_test_pred <- predict(comments_classifier, comments_test)
CrossTable(comments_test_pred, comments_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

The cross table will list how many of the predicted categories are in their correct categories. Adding diagonally and dividing will calculate the percent accuracy.

```{r}
(44+473+1142+206)/23.35
```

To make a table showing predicted categories for comments based on the Naive Bayes test, versus the actual label entered by a human observer, merge the two datasets into a single data frame. The full data frame will be too long to display; use "head" to check format, then output the full dataset to CSV for further review. By default the CSV file should be saved to the current working directory.

```{r}
predicted_category <- data.frame(comments_test_pred,comments_test_labels)
head(predicted_category)
write.csv(predicted_category, file = "PredictedCategory_Data.csv")
```

***
START BREAK POINT
***
1. The e1071 naive bayes classifier is binomial. Need to incorporate multinomial.
2. Need to add prediction output.
3. Is dropping numerical values down to categorical yes/no variables removing useful info?

***
END BREAK POINT
***

***
***
####\

##Visualization
This code generates a wordcloud of the frequency of single words in the dataset using the package "wordcloud".
```{r}
wordcloud(comments_corpus_clean, max.words = 50, random.order = FALSE)
```

Compare wordclouds between 3 groups.
```{r}
ONE <- subset(comments_raw, type == "1_Basic")
TWO <- subset(comments_raw, type == "2_Writing")
THREE <- subset(comments_raw, type == "3_Technical")
FOUR <- subset(comments_raw, type == "4_Logic")

wordcloud(ONE$text, max.words = 50, scale = c(3, 0.5))
wordcloud(TWO$text, max.words = 50, scale = c(3, 0.5))
wordcloud(THREE$text, max.words = 50, scale = c(3, 0.5))
wordcloud(FOUR$text, max.words = 50, scale = c(3, 0.5))
```

Show the 5 most frequent words in the data:
```{r}
comment_sack <- TermDocumentMatrix(comments_corpus_clean)
comment_pack <- as.matrix(comment_sack)
comment_snack <- sort(rowSums(comment_pack), decreasing = TRUE)
comment_hack <- data.frame(word = names(comment_snack), freq=comment_snack)
head(comment_hack, 10)
```

And the 5 most frequent words from each class:
```{r}
wordcloud(ONE$text, max.words = 10, scale = c(3, 0.5))
wordcloud(TWO$text, max.words = 10, scale = c(3, 0.5))
wordcloud(THREE$text, max.words = 10, scale = c(3, 0.5))
wordcloud(FOUR$text, max.words = 10, scale = c(3, 0.5))
```

####\

##Permutation Tests of Naive Bayes (NB) Method
Version 1.0 correctly identifies comment subcategories ~79% of the time when using single words, and 68.3% of the time when using 2-grams. Permutations to test:

*  Compare 2-grams, 3-grams rather than single words
*  Make pair-wise comparisons versus comparing multiple groups simultaneously. 
    +  Ex.: Basic vs. Not Basic (1_Basic vs. not 1_Basic)
*  Modify the variables within the NB analysis itself. 
    +  Laplace constant (Y/N), Cutoff value for words removed from list.
*  Do/Do not stem the terms
    +  This may remove evidence of tense
*  Remove/Leave stopwords
*  Remove/Leave capitalization
*  Remove/Leave numbers
    +  This may remove evidence of technical descriptors
*  Remove/Leave punctuation
    +  This may remove evidence of questions vs statements