### 2019-02-21 hiding/questions.rmd

# R Code
#### I cannot get numbers on CSV extracted data frames.

Detailed notes are in "Data_intake_1.rmd", starting Line 102. Any ideas?

####\

#### Why can I not append the read.csv command?
This works fine:
```
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18.csv')
head(base_data)
```

This fails:
```
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18.csv', header=FALSE)
head(base_data)
```
I get:
```	
	Error in read_csv(file = "data/coded_full_comments_dataset_Spring18.csv",  : 
  unused argument (header = FALSE)
```

-----

# Project

-----

# Github
#### Question about fork and branch. 
When I fork another repo, I get comments saying I am "#" steps ahead of it. Should I clone instead?

-----

# Archived
#### How to create files in R, then push to GitHub as a complete repo
Instructions for doing it on this website. [https://cfss.uchicago.edu/git05.html]()

-----

####\