### 2019-02-21 nextsteps.Rmd

---
title: "Next Steps"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

Qualitative coding of TA comments is informative but unsustainable. Developing the codebook then rating the first set of 11,000 comments required nearly 2 months of investigator effort. Even with the codebook, individual coders will require training to achieve any accuracy. Given 10-12,000 comments are generate EACH SEMESTER, an automated sorting method is needed.

####\