### 2019-02-23 Naive_Bayes_Pkg_test.Rmd
update

---
title: "Naive Bayes Package Test"
author: "Dan Johnson"
date: "1/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Trying out Naive Bayes Package

Source of package:
CRAN repository

```{r}
library(tidyverse)
```

```{r}
glimpse(iris)

```

```{r}
data(iris)
nb <- naive_bayes(Species ~ ., data = iris)
plot(nb)

```

```{r}
nb_kernel <- naive_bayes(x = iris[-5], y = iris[ ,5], usekernel = TRUE)
plot(nb_kernel)

```

***
BRANCH
***

```{r}
library(naivebayes)
```

```{r}
#Select rows representing the sub-groups to compare. 
comments_subsetVERB <- filter(base_dataVERB,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")

#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
comments_rawVERB <- comments_subsetVERB %>% select(23,14,15,16,17,18,19,20)

#Rename the columns.
names(comments_rawVERB)[1] <- "type"

glimpse(comments_rawVERB)

```

```{r}
data(comments_rawVERB)
nb <- naive_bayes(type ~ ., data = comments_rawVERB)
plot(nb)

```

```{r}
nb_kernel <- naive_bayes(x = iris[-5], y = iris[ ,5], usekernel = TRUE)
plot(nb_kernel)

```

***
END BRANCH
***

```{r}
X1 <- as.data.frame(matrix(sample(letters[5:9], vars * rows, TRUE),
                           ncol = vars))
nb_cat <- naive_bayes(x = X1, y = y)
nb_cat
system.time(pred2 <- predict(nb_cat, X1))

```

## Not run: 
vars <- 10
rows <- 500000
y <- sample(c("a", "b"), rows, TRUE)

# Only categorical variables
X1 <- as.data.frame(matrix(sample(letters[5:9], vars * rows, TRUE),
                           ncol = vars))
nb_cat <- naive_bayes(x = X1, y = y)
nb_cat
system.time(pred2 <- predict(nb_cat, X1))

# Only numeric variables
X2 <- as.data.frame(matrix(rnorm(vars * rows), ncol = vars))
nb_num <- naive_bayes(x = X2, y = y)
nb_num
system.time(pred2 <- predict(nb_num, X2))

## End(Not run)