### 2019-02-23 Naive_bayes_notes_5.Rmd

---
title: "Permutations of Naive Bayes"
author: "Dan Johnson"
date: "12/21/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Permutation Tests of Naive Bayes (NB) Method
Version 1.0 correctly identifies comment subcategories ~79% of the time when using single words, and 68.3% of the time when using 2-grams. Permutations to test:

Uni-grams, four categories
Lowercase=YES, remove_numbers=no, remove_stopwords=no, stemming=no, remove_whitespace=YES
Freq_cutoff=5
  Overall Discrimination: 78.6%
  Highest Discrimination: 91.4%, 3_Technical

Bi-grams, four categories
Lowercase=YES, remove_numbers=no, remove_stopwords=no, stemming=no, remove_whitespace=YES
Freq_cutoff=5
  Overall Discrimination: 73.6%
  Highest Discrimination: 82.7%, 3_Technical

Uni-grams, four categories
Lowercase=YES, remove_numbers=no, remove_stopwords=YES, stemming=no, remove_whitespace=YES
Freq_cutoff=5
  Overall Discrimination: 
  Highest Discrimination:

Bi-grams, four categories
Lowercase=YES, remove_numbers=no, remove_stopwords=YES, stemming=no, remove_whitespace=YES
Freq_cutoff=5
  Overall Discrimination: 
  Highest Discrimination:

*  Compare 2-grams, 3-grams rather than single words
*  Make pair-wise comparisons versus comparing multiple groups simultaneously. 
    +  Ex.: Basic vs. Not Basic (1_Basic vs. not 1_Basic)
*  Modify the variables within the NB analysis itself. 
    +  Laplace constant (Y/N), Cutoff value for words removed from list.
*  Do/Do not stem the terms
    +  This may remove evidence of tense
*  Remove/Leave stopwords
*  Remove/Leave capitalization
*  Remove/Leave numbers
    +  This may remove evidence of technical descriptors
*  Remove/Leave punctuation
    +  This may remove evidence of questions vs statements

##Initial Setup
```{r}
library(tidyverse)
library(tidytext)
library(tidyr)
library(tm)
library(SnowballC)
library(wordcloud)
library(e1071)
library(gmodels)
```

####\

###Initial Data Input
```{r}
#Read in TA comments from CSV file.
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')

#Select rows representing the sub-groups to compare. 
comments_subset <- filter(base_data,code.subject=="1. Basic Criteria"|code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")

#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
comments_raw <- comments_subset %>% select(23,22)

#Rename the columns.
names(comments_raw)[1] <- "type"
names(comments_raw)[2] <- "text"

#Simplify coding terms
comments_raw[,1] <- ifelse(comments_raw[,1] == "1. Basic Criteria","1_Basic", ifelse(comments_raw[,1] == "2. Writing Quality","2_Writing", ifelse(comments_raw[,1] == "3. Technical and Scientific","3_Technical", ifelse(comments_raw[,1] == "4. Logic and Thinking","4_Logic",99))))

#Change "type" element from character to a factor for analysis.
comments_raw$type <- factor(comments_raw$type)
str(comments_raw$type)
table(comments_raw$type)
```

Data set must be converted to a volative corpus using "tm" library then transformed.
```{r}
#Create the volatile coprus that contains the "text" vector from data frame.
comments_corpus <- VCorpus(VectorSource(comments_raw$text))
print(comments_corpus)

#Check out the first few messages in the new corpus, which is basically a list that can be manipulated with list operations.
inspect(comments_corpus[1:3])

#Use "as.character" function to see what a message looks like.
as.character(comments_corpus[[3]])
```

The OPTIONAL data transforms using "tm_map"" and "content_transformer" functions. Switch between "{r}" and "{}" to turn a transformation block off or on.
```{r}
#Convert to all lower case letters.
comments_corpus_clean <- tm_map(comments_corpus, content_transformer(tolower))
```

```{}
#Remove numbers. 
comments_corpus_clean <- tm_map(comments_corpus_clean, removeNumbers)
```

```{r}
#Stopword removal. 
comments_corpus_clean <- tm_map(comments_corpus_clean, removeWords, stopwords())
```

```{}
#Remove punctuation using the "removePunctuation" function. 
#This step removes evidence of questions, so may remove data.
comments_corpus_clean <- tm_map(comments_corpus_clean, removePunctuation)
```

```{}
#Stemming the text data to strip the suffix from words.  
comments_corpus_clean <- tm_map(comments_corpus_clean, stemDocument)
```

Final prep and data check.
```{r}
#Final step removes white space from the document. This is NOT optional.
comments_corpus_clean <- tm_map(comments_corpus_clean, stripWhitespace)

#Look at an example of data.
as.character((comments_corpus_clean[[3]]))
```

####\

##Modified Naive Bayes (NB) Method Using N-grams
Uses tm's NLP commands to generate either 1-, 2-, or 3-grams with the same command. The only change needed is "#" in "(words(x), #)" below.
```{r}
NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
}

comments_dtm_2gram <- DocumentTermMatrix(comments_corpus_clean, control=list(tokenize = NLP_Tokenizer))
comments_dtm_2gram_tidy <- tidy(comments_dtm_2gram)
inspect(comments_dtm_2gram)
```

####\

###Data Preparation
Split data into 75% training and 25% testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data.

```{r}
.75 * 9340 #Number of rows in the dataset; product is #rows for training
.25 * 9340 #Product is #rows for testing set
```

This code assumes comments are random. **Probably want to try randomizing them.**
```{r}
comments_dtm_train <- comments_dtm_2gram[1:7005, ]
comments_dtm_test <- comments_dtm_2gram[7006:9340, ]
```

Save vectors labeling rows in the training and testing vectors
```{r}
comments_train_labels <- comments_raw[1:7005, ]$type
comments_test_labels <- comments_raw[7006:9340,]$type
```

Make sure that the proportion of each sub-category is similar in the training and testing data set.
```{r}
prop.table(table(comments_train_labels))
prop.table(table(comments_test_labels))
```

####\

###Preparation for Naive Bayes
Remove words from the matrix that appear less than 5 times.
```{r}
comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
str(comments_freq_words)
```

Limit our Document Term Matrix to only include words in the sms_freq_vector. We want all the rows, but we want to limit the column to these words in the frequency vector.
```{r}
comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
```

The naive bayes classifier works with categorical features, so we need to convert the matrix to "yes" and "no" categorical variables. To do this we'll build a convert_counts function and apply it to our data.
```{r}
convert_counts2 <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}
```

This replaces values greater than 0 with yes, and values not greater than 0 with no. Let's apply it to our data.
```{r}
comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
```

The resulting matrixes will be character type, with cells indicating "yes" or "no" if the word represented by the column appears in the message represented by the row.

####\

###Train Model, Predict, Evaluate
Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group ONE, TWO, THREE, or FOUR. Evaluate the prediction with the actual data using a crosstable from the gmodels package.
```{r}
comments_classifier <- naiveBayes(comments_train, comments_train_labels)
comments_test_pred <- predict(comments_classifier, comments_test)
CrossTable(comments_test_pred, comments_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

```{r}
(44+375+1116+183)/23.35
```

***
***