### 2019-02-23 docs/hiding/reference_tables.Rmd

---
title: "Reference Tables"
author: "Dan Johnson"
date: "2/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#TA Comment Categories
Labels in **comment.subject** are routinely recoded as:

0_null
1_basic
2_writing
3_technical
4_logic
5_exposition
6_misconduct
7_administrative
8_nobasis
9_split
10_context
11_flag
12_narrative
13_sciname

Labels in **comment.structure** are routinely recoded as:

0_null
1_pointer
2_copyedit
3_general
4_specific
5_rationale
6_holistic
7_idiomatic
8_nobasis