### 2019-02-23 reference_tables.Rmd

---
title: "Automated Classification of TA Comments on Student Writing"
author: "Dan Johnson"
date: "2/3/2019"
output: html_document
---

```{r, child="_setup.Rmd"}
```

#TA Comment Categories
Labels in **comment.subject** are routinely recoded as:

*  0_null
*  1_basic
*  2_writing
*  3_technical
*  4_logic
*  5_exposition
*  6_misconduct
*  7_administrative
*  8_nobasis
*  9_split
*  10_context
*  11_flag
*  12_narrative
*  13_sciname

Labels in **comment.structure** are routinely recoded as:

*  0_null
*  1_pointer
*  2_copyedit
*  3_general
*  4_specific
*  5_rationale
*  6_holistic
*  7_idiomatic
*  8_nobasis