### 2019-02-24 build.Rmd
update

---
title: "The Build"
date: "2/22/2019"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

#Overview
This page links to sub-pages that describe the build process for various components used in the overall project. Follow the links for details.

Extracting TAs' comments (*draft*): https://adanieljohnson.github.io/default_website/docxcommentextraction.html

Codebook for classifying TAs' comments: https://adanieljohnson.github.io/default_website/codebook.html

Structure of development dataset; data import & validation: https://adanieljohnson.github.io/default_website/dataimport.html

Current version of the Naive Bayes classifier: https://adanieljohnson.github.io/default_website/NB_2r2c.html

Expanding the Naive Bayes classifier to an Ensemble Model (*draft*): https://adanieljohnson.github.io/default_website/Ensemble_Notes.html

####\