### 2019-02-25 hiding/Latent_dirichlet_allocation_notes.Rmd

---
title: "Latent Dirichlet Analysis v1"
author: "Dan Johnson"
date: "12/14/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Sources:
Primary
https://eight2late.wordpress.com/2015/09/29/a-gentle-introduction-to-topic-modeling-using-r/

```{r}
#load text mining library
library(tm)

#set working directory (modify path as needed)
setwd("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/ta_comments_FLC/data/textmining/")

#load files into corpus
#get listing of .txt files in directory
filenames <- list.files(getwd(),pattern="*.txt")

#read files into a character vector
files <- lapply(filenames,readLines)

#create corpus from vector
docs <- Corpus(VectorSource(files))

#inspect a particular document in corpus
writeLines(as.character(docs[[30]]))

#start preprocessing
#Transform to lower case
docs <-tm_map(docs,content_transformer(tolower))

#remove potentially problematic symbols
toSpace <- content_transformer(function(x, pattern) { return (gsub(pattern, " ", x))})
docs <- tm_map(docs, toSpace, "-")
docs <- tm_map(docs, toSpace, "’")
docs <- tm_map(docs, toSpace, "‘")
docs <- tm_map(docs, toSpace, "•")
docs <- tm_map(docs, toSpace, "“")
docs <- tm_map(docs, toSpace, "”")

#remove punctuation
docs <- tm_map(docs, removePunctuation)
#Strip digits
docs <- tm_map(docs, removeNumbers)
#remove stopwords
docs <- tm_map(docs, removeWords, stopwords("english"))
#remove whitespace
docs <- tm_map(docs, stripWhitespace)
#Good practice to check every now and then
writeLines(as.character(docs[[30]]))
#Stem document
docs <- tm_map(docs,stemDocument)

#fix up 1) differences between us and aussie english 2) general errors
docs <- tm_map(docs, content_transformer(gsub),
pattern = "organiz", replacement = "organ")
docs <- tm_map(docs, content_transformer(gsub),
pattern = "organis", replacement = "organ")
docs <- tm_map(docs, content_transformer(gsub),
pattern = "andgovern", replacement = "govern")
docs <- tm_map(docs, content_transformer(gsub),
pattern = "inenterpris", replacement = "enterpris")
docs <- tm_map(docs, content_transformer(gsub),
pattern = "team-", replacement = "team")

#define and eliminate all custom stopwords
myStopwords <- c("can", "say","one","way","use",
"also","howev","tell","will",
"much","need","take","tend","even",
"like","particular","rather","said",
"get","well","make","ask","come","end",
"first","two","help","often","may",
"might","see","someth","thing","point",
"post","look","right","now","think","‘ve ",
"‘re ","anoth","put","set","new","good",
"want","sure","kind","larg","yes,","day","etc",
"quit","sinc","attempt","lack","seen","awar",
"littl","ever","moreov","though","found","abl",
"enough","far","earli","away","achiev","draw",
"last","never","brief","bit","entir","brief",
"great","lot")
docs <- tm_map(docs, removeWords, myStopwords)

#inspect a document as a check
writeLines(as.character(docs[[30]]))

#Create document-term matrix
dtm <- DocumentTermMatrix(docs)

#convert rownames to filenames
rownames(dtm) <- filenames

#collapse matrix by summing over columns
freq <- colSums(as.matrix(dtm))

#length should be total number of terms
length(freq)

#create sort order (descending)
ord <- order(freq,decreasing=TRUE)

#List all terms in decreasing order of freq and write to disk
freq[ord]
write.csv(freq[ord],"word_freq.csv")
```

##Part 2

```{r}
#load topic models library
library(topicmodels)

#set working directory (modify path as needed)
setwd("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/ta_comments_FLC/data/textmining/")

#Set parameters for Gibbs sampling
burnin <- 4000
iter <- 2000
thin <- 500
seed <-list(2003,5,63,100001,765)
nstart <- 5
best <- TRUE

#Number of topics
k <- 5

#That done, we can now do the actual work – run the topic modelling algorithm on our corpus. Here is the code:

#Run LDA using Gibbs sampling
ldaOut <-LDA(dtm,k, method="Gibbs", control=list(nstart=nstart, seed = seed, best=best, burnin = burnin, iter = iter, thin=thin))

#write out results
#docs to topics
ldaOut.topics <- as.matrix(topics(ldaOut))
write.csv(ldaOut.topics,file=paste("LDAGibbs",k,"DocsToTopics.csv"))

#top 6 terms in each topic
ldaOut.terms <- as.matrix(terms(ldaOut,6))
write.csv(ldaOut.terms,file=paste("LDAGibbs",k,"TopicsToTerms.csv"))

#probabilities associated with each topic assignment
topicProbabilities <- as.data.frame(ldaOut@gamma)
write.csv(topicProbabilities,file=paste("LDAGibbs",k,"TopicProbabilities.csv"))

#Find relative importance of top 2 topics
topic1ToTopic2 <- lapply(1:nrow(dtm),function(x)
sort(topicProbabilities[x,])[k]/sort(topicProbabilities[x,])[k-1])

#Find relative importance of second and third most important topics
topic2ToTopic3 <- lapply(1:nrow(dtm),function(x)
sort(topicProbabilities[x,])[k-1]/sort(topicProbabilities[x,])[k-2])

#write to file
write.csv(topic1ToTopic2,file=paste("LDAGibbs",k,"Topic1ToTopic2.csv"))
write.csv(topic2ToTopic3,file=paste("LDAGibbs",k,"Topic2ToTopic3.csv"))
```

***
***

##Other Examples of LDA Procedures
https://rpubs.com/Junny31/318845
https://rstudio-pubs-static.s3.amazonaws.com/266565_171416f6c4be464fb11f7d8200c0b8f7.html

***
***

##Another Approach Comparing LDA (unsupervised) and tf-idf (supervised)

https://www.kaggle.com/rtatman/nlp-in-r-topic-modelling
>Latent Dirichlet (said like deer-ih-CLAY) allocation, more commonly shortened to LDA, is one unsupervised way of doing topic modeling. In machine learning, a problem is unsupervised if you don't have labeled examples of the thing you're trying to figure out.

>LDA rests on the assumption that each document is going to have some subset of topics in it, and that each topic is associated with a certain subset of words. Imagine, for example, that you have three text ads:

    One is for Chow-Chow Crunch, a new dog food
    One is for Danny's Downhome Diner, a new restaurant
    One is for Fifi's Chic Puppy Boutique, a store that sells dog accessories

>Both the diner and the dog food ad will probably have words associated with the topic of food in them, like "tasty", "filling" or "delicious". Both the dog food and the pet store will probably have words associated with the topic of pets, like "canine", "pet" or "owner". And, because they're all ads, it likely that all three documents will have words associated with advertising in them, like "sale", "price" or "special".

>The tricky part is that words can also belong to more than one topic. I just said that "special" could be associated with advertising... but it can also be associated with food, especially in terms like "blue plate special" or "chef's special".

>What LDA outputs, then, is two estimates:

    An estimate of how much each topic contributes to each document
    An estimate of how much each word contributes to each topic

>For exploratory analysis, especially if you have a large number of documents, the second estimate is usually more interesting. In order to make our lives easier (since we're going to ran an LDA several times) let's write a function that takes a text column from a data frame and returns a plot of the most informative words for a given number of topics.

```{r}
# function to get & plot the most informative terms by a specificed number
# of topics, using LDA
top_terms_by_topic_LDA <- function(input_text, # should be a columm from a dataframe
                                   plot = T, # return a plot? TRUE by defult
                                   number_of_topics = 4) # number of topics (4 by default)
{    
    # create a corpus (type of object expected by tm) and document term matrix
    Corpus <- Corpus(VectorSource(input_text)) # make a corpus object
    DTM <- DocumentTermMatrix(Corpus) # get the count of words/document

    # remove any empty rows in our document term matrix (if there are any 
    # we'll get an error when we try to run our LDA)
    unique_indexes <- unique(DTM$i) # get the index of each unique value
    DTM <- DTM[unique_indexes,] # get a subset of only those indexes

    # preform LDA & get the words/topic in a tidy text format
    lda <- LDA(DTM, k = number_of_topics, control = list(seed = 1234))
    topics <- tidy(lda, matrix = "beta")

    # get the top ten terms for each topic
    top_terms <- topics  %>% # take the topics data frame and..
      group_by(topic) %>% # treat each topic as a different group
      top_n(10, beta) %>% # get the top 10 most informative words
      ungroup() %>% # ungroup
      arrange(topic, -beta) # arrange words in descending informativeness

    # if the user asks for a plot (TRUE by default)
    if(plot == T){
        # plot the top ten terms for each topic in order
        top_terms %>% # take the top terms
          mutate(term = reorder(term, beta)) %>% # sort terms by beta value 
          ggplot(aes(term, beta, fill = factor(topic))) + # plot beta by theme
          geom_col(show.legend = FALSE) + # as a bar plot
          facet_wrap(~ topic, scales = "free") + # which each topic in a seperate plot
          labs(x = NULL, y = "Beta") + # no x label, change y label 
          coord_flip() # turn bars sideways
    }else{ 
        # if the user does not request a plot
        # return a list of sorted terms instead
        return(top_terms)
    }
}
```