### 2019-02-25 docs/hiding/Check_in_Jan_14.Rmd
Web site update

---
title: "Jan 14 Check-In"
author: "Dan Johnson"
date: "1/29/19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#New ideas From FLC Meeting #5

Check my data for correlations.
	"cor" blows up on missing data
	Try "use" pairwise complete

*  Use R profile for storing API tokens

```{r}
file.edit(file.path("~", ".Rprofile")) # edit .Rprofile in HOME
file.edit(".Rprofile") # edit project specific .Rprofile

# There is information about the options you can set via
# help("Rprofile")
```

Tools to consider:

*  UDPipe for syntactic parsing
*  Stylometry library
*  QualtRics package for connecting to API (no longer supported by CRAN)
*  By hand exploration followed up with computational exploration

Reference sources:

*  R Open Sci

To Do List:

*  Need to add some data intermediates to my work to help others visualize.
*  Taskviews on CRAN
*  Look at lemmatization, not stemming

Look into Bayes model commands:

*  prob
*  predict

Packages to evaluate

*  caret
*  skimR
*  parsnip
*  recipe
*  tidymodels
*  UDPipe
*  package = usethis

##What Is Working

Solved the data confidentiality problem. Turned out to be easier and more secure to do outside of R.

Manipulating the column vectors.
  Suggested: I need to recode some values to simplify. 
Second, in this same correlation analysis, I would recommend recoding the values in the `code.subject` column. You have some type of error there and I tend to avoid calling values such as "2. Writing Quality" as they tend make code more difficult to visually parse (and therefore debug) and it's also easy to get something wrong with the value (even though it seems that you've directly copied and pasted the values to double check).

For recoding comment labels, this is what I plan to use:

Recoding Comment.subject
0_null
1_basic
2_writing
3_technical
4_logic
5_exposition
6_misconduct
7_administrative
8_nobasis
9_split
10_context
11_flag
12_narrative
13_sciname

Recoding Comment.structure
0_null
1_pointer
2_copyedit
3_general
4_specific
5_rationale
6_holistic
7_idiomatic
8_nobasis

Calculation and plotting correlations and log-odds pairs is working (except for broken parts that recoding should solve hopefully.)

Suggested: Check my data for small errors that break correlations.
	"cor" blows up on missing data
	Try "use" pairwise complete

Tokenizer and terms cleanup seems to work well. (Still struggling with understanding how data get transformed in certain copied code blocks.)
  Suggested: Look into lemmatization, not stemming

Have an operational categorical naive Bayes workflow that can perform analysis using 1-, 2-, or 3-grams.

## Coding & Interpretation Questions

1.  Randomization. Does the straight split I used bias the data? 
  Comment: YES, so instead use the following code to pick more randomized dataset.

```{r}
library(dplyr)
set.seed(1234) #randomization will be same set each time
train <- sample_frac(mtcars, .8) #sets how many are in random training set
test <- setdiff(mtcars, train) #chooses non-training for test set
```

Suggestions:
  Try: 
    Leave one out cross-validation
    K-folds cross-validation

I do have a few things for you to consider before we do get together. The first thing I suggest you do, which shouldn't be too much of a problem to create, is to provide the output of some of your datasets (particularly thinking about the contrast sets that you use in the correlation analysis). A quick `glimpse(data)` printed out would help me get a better sense of the structure you are using as input and output at each stage of your analysis.

Third thing that comes to mind has to do with the Naive Bayes analysis. As you've seen, naive bayes function from the `e1071` package only works with categorical features. For many NLP tasks this oversimplifies features making the appearance of a word a 'yes' or 'no' decision, which reduces information concerning the frequency of occurrence of words. It's been a while since I've used naive bayes in my own work, but last time I did, I ended up writing my own algorithm to deal with non-categorical features. I'm not altogether happy with the algorithm I coded, but when we meet we can talk about whether there are other package options out there now, or whether my function might be of some help to you.

Anyway, things are looking good. Keep up the good work and great documentation!

    FLC - Data Analysis (AB)

My project is still taking shape (and isn't yet really research, per se), but a few things I've been working on:

    Project Updates - Updating some content on my Project page
    New Page - Learning Analytics. This is a proposal I have had accepted for an April conference, and I am hoping our group might help me develop some of my thinking on the issue.
    Course Notes - I haven't kept up with note taking in this space, but added some of the detail from my work during Session 4. I am looking forward to engaging more actively with this space during our actual sessions, and I am interested in hearing about others' thoughts on this process.

While there are many more updates I hope to make this term, I wanted to share some of this work. You might also check out Dan's page (as his is more robust than mine).

To-Do

    Prepare to present on your progress at Monday's Session
    Update your website (or have a plan to do so).
    Raise questions (technical or otherwise) in the TLC discussion board space.

I am looking forward to seeing everyone on Monday (assuming we aren't snowed out).

2.  How to ID, prune unneeded packages.

3.  When I join columns in correlation analysis, when a word is MISSING from **any subset** the data for that word is pruned from the full dataset. I am losing the unique words this way. Need to figure out how to join without pruning.

6.  Broken correlation analyses. I have a problem buried in the coded data that is throwing an error in correlations. Jerid recommended recoding the values in the `code.subject` column.

## Bigger Questions
Python/Scython porting

After posting an initial naive Bayes trial, Jerid wrote:
>Naive bayes function from `e1071` package only works with categorical features. For many NLP tasks this oversimplifies features making the appearance of a word a 'yes' or 'no' decision.

I spent time looking at various methods as part of background research. The summary is posted elsewhere. Question is whether I should focus on:

*  Continued refinement of NB, or exploring other supervised models? 
    +  How deep to go into different methods?
    +  When to drop one model, move to another?
*  Using my existing code-book to reverse engineer a pure rules based method instead?
    +  Faster for me, but what about others?
*  Extending previous item, create a HYBRID of statistical and rules-based?
    +  Words to evaluate are pre-selected limited vocabulary (see codebook for lists)
    +  Words are scored present/absent in a comment, then comment sorted based on most likely.
*  Step back and rethink the categories using LDA, etc.? 
    +  Seems counter-productive given I have a model built already.