### 2019-02-25 litreview.Rmd
correction

---
title: "Literature Review"
date: "2/21/2019"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

#Overview
The supporting literature for this project is broad and convoluted. Rather than try to compress it to a single page, I have provided links out to sub-pages that describe various background topics. Follow the links for details.

Fundamental Principles and Practices of Technical Writing Training:  https://adanieljohnson.github.io/default_website/pagename.html

Primer on Text Classification Strategies:  https://adanieljohnson.github.io/default_website/Text_Classifier_Models.html

####\