### 2019-02-25 build.Rmd

---
title: "The Build"
date: "2/22/2019"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

This page links to sub-pages describing how various components used in the overall project were constructed. Follow the links for details.

Extracting TAs' Comments (***Draft***): https://adanieljohnson.github.io/default_website/docxcommentextraction.html

Codebook for Classifying TAs' Comments: https://adanieljohnson.github.io/default_website/codebook.html

Structure of Development Dataset; Data Import & Validation: https://adanieljohnson.github.io/default_website/dataimport.html

Initial Exploration of the Dataset: https://adanieljohnson.github.io/default_website/initial_exploration.html

Current Version of the Naive Bayes Comment Classifier: https://adanieljohnson.github.io/default_website/NB_2r2c.html

Embedding the Naive Bayes Classifier in an Ensemble Model (***Draft***): https://adanieljohnson.github.io/default_website/Ensemble_Notes.html

####\