### 2019-02-25 litreview.Rmd

---
title: "Literature Review"
date: "2/21/2019"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

#Overview
This project brings together several traditionally separate fields of research:

*  General learning theories
*  Educational research specific to STEM fields
*  Writing in the disciplines, and technical writing instruction
*  Text analysis and classification

Instead of trying to weave a coherent one-page narrative of these very different fields, I have provided links out to sub-pages that describe various background topics separately. Follow the links for details.

General Learning Theories Relevant to the Current Challenge:
https://adanieljohnson.github.io/default_website/pagename.html

Educational Research Specific to STEM fields
https://adanieljohnson.github.io/default_website/pagename.html

Fundamental Principles and Practices of Technical Writing Training:  https://adanieljohnson.github.io/default_website/pagename.html

Primer on Text Classification Strategies; Feature Selection and Engineering:  https://adanieljohnson.github.io/default_website/Text_Classifier_Models.html

>Sub-pages:  
>1.  Topic categorization and modeling (https://adanieljohnson.github.io/default_website/Topic_categorization.html)  
>2.  Selecting an appropriate text classifier method (https://adanieljohnson.github.io/default_website/Choosing_classifier.html)  
>3.  Evaluating the outcomes, and improving the classifier (https://adanieljohnson.github.io/default_website/Improving_classifier.html)

####\