### 2019-02-25 Help.rmd
updates

Items to get help on:

#Index:
	Site map

#Background:

#Lit review:
	How to branch to Text Classification Primer, sub-pages look okay?
		Text_classifier_models	= 	Text Classification Primer
		Topic_categorization = Topic Categorization and Modeling Methods
		Choosing_classifier = Choosing a Text Classifier Method
		Classifier_comparison = Designing a Classifier Evaluation Workflow 
		Improving_classifier = Strategies for Improving Text Classifiers

#The Build:
	df
	  df

#Data Exploration:
	Help with aligning tables and keeping the NA values

#Findings:
	How to graph the different comparisons of NB values from different sources on same plot.

#Next Steps:
	Df
	  Df

#Etc.:
	Using a Bibliography in this format
	Can menus have sub-pages
	Adding automatic updates to top dates
	Moving top dates

Using NLP and text data mining