### 2019-02-25 notes.rmd
update

---
title: "Working Notes"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

#Items to get help on:

##Index:
	Site map

##Background:
	Df
	  Df

##Lit review:
	How to branch to Text Classification Primer, sub-pages look okay?
		Text_classifier_models	= 	Text Classification Primer
		Topic_categorization = Topic Categorization and Modeling Methods
		Choosing_classifier = Choosing a Text Classifier Method
		Classifier_comparison = Designing a Classifier Evaluation Workflow 
		Improving_classifier = Strategies for Improving Text Classifiers

##The Build:
	Df
	  Df

##Data Exploration:
	Help with aligning tables and keeping the NA values

##Findings:
	How to graph the different comparisons of NB values from different sources on same plot.

##Next Steps:
	Still in draft stage
	  Df

##Etc.:
	Using a Bibliography in this format?
	Can menus have sub-pages?
	Adding automatic updates to top dates?
	Moving top dates?
	Fix GitHub auto-upload

##Check These Out:
  Upload 'devtools::install_github("WFU-TLC/analyzr")'
  Using Mike's Metadata packager to annotate our released dataset
  Try using ```mutate(datacolumn=str_count(utterance_clean, "\\b(um|u=m)\\b"))``` to extract a string character and count.

####\