### 2019-02-25 litreview.Rmd
update

---
title: "Literature Review"
date: "`r format(Sys.time(), '%b %d, %Y')`"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

#Overview
This project brings together several traditionally separate fields of research:

*  General learning theories
*  Educational research specific to STEM fields
*  Writing in the disciplines, and technical writing instruction
*  Text analysis and classification

Instead of trying to weave a coherent one-page narrative of these very different fields, I have provided links out to sub-pages for each of the various background topics. Follow the links for details.

General Learning Theories Relevant to the Current Challenge:
https://adanieljohnson.github.io/default_website/pagename.html

STEM Educational Research:
https://adanieljohnson.github.io/default_website/pagename.html

Foundations of Technical Writing Training:  https://adanieljohnson.github.io/default_website/pagename.html

Text Classification 1 - Feature Selection and Engineering:  https://adanieljohnson.github.io/default_website/Text_Classifier_Models.html

Text Classification 2 - Topic categorization and modeling (https://adanieljohnson.github.io/default_website/Topic_categorization.html)

Text Classification 3 - Selecting an appropriate text classifier method (https://adanieljohnson.github.io/default_website/Choosing_classifier.html)

Text Classification 4 - Evaluating the outcomes, and improving the classifier (https://adanieljohnson.github.io/default_website/Improving_classifier.html)

####\