### 2019-03-13 build.Rmd

---
title: "Project Build-Out"
date: "`r format(Sys.time(), '%b %d, %Y')`"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

This page links to sub-pages describing construction of the various components used in the overall project. Progress on each component is in parentheses.

####\

####Extracting TAs' Comments 
*(Early coding stage)*  
Protocol used to extract TA comments from student lab reports.  
https://adanieljohnson.github.io/default_website/docxcommentextraction.html

####Structure of Development Dataset; Data Import & Validation 
*(Completed)*  
Organization of the initial dataset, and how data were imported in R and validated.  
https://adanieljohnson.github.io/default_website/dataimport.html

####Codebook for Classifying TAs' Comments 
*(Completed)*  
Describes the process for developing the codes used to classify comments.  
https://adanieljohnson.github.io/default_website/codebook.html

####Initial Exploration of the Dataset 
*(Partly completed)*  
What is the distribution of words used in the TA comments?  
https://adanieljohnson.github.io/default_website/initial_exploration.html

####Current Version of the Naive Bayes Comment Classifier 
*(Completed)*  
Code for the most recent version of the optimized NB classifier.   
https://adanieljohnson.github.io/default_website/NB_2r2c.html

####Embedding the Naive Bayes Classifier in an Ensemble Model  
*(Planning stage)*
The NB method alone does not classify TA comments with sufficient accuracy for the "Structure" category. This page outlines a vertically stacked "ensemble" classifier that uses pattern matching and other pre-processing steps to improve NB accuracy.
https://adanieljohnson.github.io/default_website/Ensemble_Notes.html

####\  
***