### 2019-05-12 N_gram_exploration.rmd

---
title: "N-gram Frequency Analysis"
date: "`r format(Sys.time(), '%b %d, %Y')`"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

Another comparison looks at frequencies of N-grams between different sub-categories. This page describes the process used for initial exploration of the TA comments dataset.

####\

#N-Grams Mapping 
##Are Certain Terms and Phrases Unique to One Sub-Category?
One of our earliest observations was that the n-gram "basic criteria" only appeared in TA comments from three different Subject sub-categories. Of these, the phrase made up <0.5% of comments in two categories, but in the third category, "basic criteria" made up >5% of all comments. This suggests that the "basic criteria" n-gram is essentially diagnostic for that sub-category. This suggests other terms and n-grams that are eliminated in the previous two methods may be equally diagnostic for one sub-category.

The block below describes a pseudo-code approach.
```{}
Import comments
Tokenize the comments as 1-, 2, 3-grams
Add counts within subset, calculate frequency
Split by subcategory
Align 4 subcategories, highest to lowest

```

####\  
```{r}
# Call in tidyverse, other required libraries.
library(tidyverse)
library(tidytext)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)

#Read in CSV as a starting dataframe named "base_data". 
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

```{r}
#Tokenize phrases as N-grams.

base_data_tokenized2n <- base_data %>% 
  unnest_tokens(ngram, ta.comment, token="ngrams", n=2)

sortedwords.comments.all2n <- base_data_tokenized2n %>%
  count(ngram, sort = TRUE) %>%
  ungroup()

total_words.all2n <- sortedwords.comments.all2n %>% 
  summarize(total = sum(n))

sortedwords.comments.all.fraction2n <- sortedwords.comments.all2n %>% as_tibble() %>% mutate(
  all.fraction = (n*100)/total_words.all2n$total
  )

#Select SUBSETs of tokenized data based on code.subject: 
base_data_tokenized.basic2n <- subset(base_data_tokenized2n, code.subject == "1_basic", select = (1:28))
base_data_tokenized.writing2n <- subset(base_data_tokenized2n, code.subject == "2_writing", select = (1:28))
base_data_tokenized.technical2n <- subset(base_data_tokenized2n, code.subject == "3_technical", select = (1:28))
base_data_tokenized.logic2n <- subset(base_data_tokenized2n, code.subject == "4_logic", select = (1:28))

#For each subset, filter stopwords, count words and total, then calculate frequencies.
sortedwords.basic2n <- base_data_tokenized.basic2n %>%
  count(ngram, sort = TRUE) %>%
  ungroup()

#Calculate the number of words in total in the subset
total_words.basic2n <- sortedwords.basic2n %>% 
  summarize(total = sum(n))

#Mutate table to calculate then append fractional word frequency in the text.
sortedwords.basic.fraction2n <- sortedwords.basic2n %>% as_tibble() %>% mutate(
  basic.fraction2n = (n*100)/total_words.basic2n$total
)

#Repeats for subset 2
#For each subset, filter stopwords, count words and total, then calculate frequencies.
sortedwords.writing2n <- base_data_tokenized.writing2n %>%
  count(ngram, sort = TRUE) %>%
  ungroup()

#Calculate the number of words in total in the subset
total_words.writing2n <- sortedwords.writing2n %>% 
  summarize(total = sum(n))

#Mutate table to calculate then append fractional word frequency in the text.
sortedwords.writing.fraction2n <- sortedwords.writing2n %>% as_tibble() %>% mutate(
  writing.fraction2n = (n*100)/total_words.writing2n$total
)

#Repeats for subset 3
#For each subset, filter stopwords, count words and total, then calculate frequencies.
sortedwords.technical2n <- base_data_tokenized.technical2n %>%
  count(ngram, sort = TRUE) %>%
  ungroup()

#Calculate the number of words in total in the subset
total_words.technical2n <- sortedwords.technical2n %>% 
  summarize(total = sum(n))

#Mutate table to calculate then append fractional word frequency in the text.
sortedwords.technical.fraction2n <- sortedwords.technical2n %>% as_tibble() %>% mutate(
  technical.fraction2n = (n*100)/total_words.technical2n$total
)

#Repeats for subset 4
#For each subset, filter stopwords, count words and total, then calculate frequencies.
sortedwords.logic2n <- base_data_tokenized.logic2n %>%
  count(ngram, sort = TRUE) %>%
  ungroup()

#Calculate the number of words in total in the subset
total_words.logic2n <- sortedwords.logic2n %>% 
  summarize(total = sum(n))

#Mutate table to calculate then append fractional word frequency in the text.
sortedwords.logic.fraction2n <- sortedwords.logic2n %>% as_tibble() %>% mutate(
  logic.fraction2n = (n*100)/total_words.logic2n$total
)
```

####\  
***
*Work Cited*