### 2019-05-13 build.Rmd
update

---
title: "Project Build-Out"
date: "`r format(Sys.time(), '%b %d, %Y')`"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

This page links to sub-pages describing construction of the various components used in the overall project. Progress on each component is in parentheses.

####\

####Extracting TAs' Comments 
*(Draft)*  
Fully R-based protocol for extracting TA comments from student lab reports. Will replace current process using a mix of Linux and Excel scripts.  
https://adanieljohnson.github.io/default_website/docxcommentextraction.html

####Structure of Development Dataset; Data Import & Validation 
Organization of the initial dataset, and how data were imported in R and validated.  
https://adanieljohnson.github.io/default_website/dataimport.html

####Codebook for Classifying TAs' Comments 
Describes the process for developing the codes used to classify comments.  
https://adanieljohnson.github.io/default_website/codebook.html

####Initial Exploration of Dataset 
What is the distribution of words used in the TA comments?  
https://adanieljohnson.github.io/default_website/initial_exploration.html

####Naive Bayes Comment Classifier 
Code for the most recent version of the optimized NB classifier.   
https://adanieljohnson.github.io/default_website/NB_2r2c.html

####Embedding Naive Bayes Classifier in an Ensemble Model  
Outlines a strategy for combining pattern matching and other pre-processing steps with Naive Bayes Classifier to improve accuracy.
https://adanieljohnson.github.io/default_website/Ensemble_Notes.html

####\